---
title: "p8105_hw3_jg4197"
author: "Jin Ge"
date: "10/5/2019"
output: github_document
---

## Exploration of instacart dataset

```{r problem 1, message = FALSE}
# open the packages and dataset
library(p8105.datasets)
library(tidyverse)
data("instacart")

# get knowledge of each variable
names(instacart)

# the number of aisles
instacart %>% 
  summarise(n_unique = n_distinct(aisle_id))
## there are 134 different aisles 

# the aisle with the most items ordered
instacart %>% 
  count(aisle) %>% 
  arrange(n = desc(n)) %>% 
  head(1)
## fresh vegetables have the most ordered items with n = 150609



# make a plot showing the number of items vs aisle

## create a data frame containing aisle names information
aisle_dat <- instacart %>% 
  select(aisle_id, aisle) %>% 
  distinct(aisle, .keep_all = TRUE) %>% 
  arrange(aisle_id)

## pic a scatterplot
instacart %>% 
  count(aisle_id) %>% 
  left_join(aisle_dat, by = "aisle_id") %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle_id, y = n)) +
  geom_point(aes(color = aisle), 
             alpha = .5) +
  labs(
    title = "The plot of the number of aisles with above 10000 items ordered",
    x = "Aisle Code",
    y = "The Number of Items Ordered(10^4)"
  ) +
  scale_x_continuous(
    breaks = c(0, 50, 100, 135)
  ) + 
  scale_y_continuous(
    breaks = c(40000, 80000, 120000),
    labels = c("4", "8", "12")
  ) +
  theme(legend.position = "bottom")
## x-aixs showing the different aisles and y-axis showing the count of each aisle. Two products are evident from other products on high number of orders in this type of products.


# a table with 3 most popular items in each aisle
instacart %>% 
  select(aisle, product_name) %>% 
  filter(aisle %in% c("baking ingredients", 
                      "dog food care", 
                      "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(count = n()) %>%
  mutate(rank = order(
    order(count, product_name, decreasing = TRUE)
  )) %>% 
  filter(rank %in% c("1", "2", "3")) %>% 
  pivot_wider(aisle:count,
              names_from = "product_name",
              values_from = "count", 
              values_fill = list(count = 0)) %>% 
  knitr::kable()
## Since each aisle sends different items, the missing value should be zero meaning that this item is not sent by this aisle


# a table with mean hour at two items ordered
instacart %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  mutate(order_dow = recode(order_dow, `0` = "Mon",
                            `1` = "Tues",
                            `2` = "Wed",
                            `3` = "Thur",
                            `4` = "Fri",
                            `5` = "Sat",
                            `6` = "Sun")) %>% 
  filter(product_name %in% c("Pink Lady Apples", 
                             "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour_day = mean(order_hour_of_day)) %>% 
  pivot_wider(product_name:mean_hour_day,
              names_from = "order_dow",
              values_from = "mean_hour_day") %>% 
  knitr::kable()

## this is a 2*7 table showing the mean hour in each day of a week

```
The size of *instacart* dataset is `r dim(instacart)` and the column types are `r class(instacart)`. The variable names are `r names(instacart)`. The key variables in this questions are _aisle_ showing different passages conveying different types of products, _product_name_ showing the product types bought by people, _order_hour_of_day_ showing the time of the order, _order_dow_ showing the day of the week. _reordered_ is a binary variable to describle whether this product is reordered by this customer.



## Exploration of BRFSS dataset

```{r problem 2, message = FALSE}
# clean the data 
brfss <- p8105.datasets::brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good",
                         "Good", "Fair", "Poor")) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good",
                                                "Very good", "Excellent")
                           ))
head(brfss, 3)


# the states observed at 7 or more location in 2002
brfss %>% 
  select(year, locationabbr, locationdesc) %>% 
  filter(year == "2002") %>% 
  group_by(locationabbr) %>% 
  summarise(n_place = n()) %>% 
  filter(n_place >= 7) %>% 
  pull(locationabbr)

# the states observed at 7 or more location in 2010
brfss %>% 
  select(year, locationabbr, locationdesc) %>% 
  filter(year == "2010") %>% 
  group_by(locationabbr) %>% 
  summarise(n_place = n()) %>% 
  filter(n_place >= 7) %>% 
  pull(locationabbr)

# construct a new dataset and make a plot
brfss %>% 
  filter(response == "Excellent") %>% 
  select(year, state = locationabbr, data_value) %>% 
  group_by(year, state) %>% 
  summarise(mean_data_value = mean(data_value)) %>% 
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_point(alpha = .5) + geom_line(alpha = .8) + theme_classic() +
  scale_x_continuous(
    breaks = c(2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010)
  ) +
  labs(
    title = "Average Excellent Health Data Across Years in Each State",
    x = "Year",
    y = "Average values of excellent health",
    caption = "Data from BRFSS"
  ) 

## The plot shows the change of each state's number of excellent heath report across year

# two-panel plot for NY in 2006 and 2010
brfss %>% 
  filter(year %in% c("2006", "2010"),
         locationabbr == "NY") %>% 
  select(year, response, locationdesc, data_value) %>% 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_boxplot(aes(fill = response), alpha = .5) +
  facet_grid(~year) +
  labs(
    title = "Health Conditon in New York in 2006 and 2010",
    x = "Response level", 
    y = "Health data values",
    caption = "Data from BRFSS"
  )
## This plot shows the values about health behavior reporting in five ersponse levels in two years
```


## exploration of a dataset about accelerometers

```{r problem 3, message = FALSE}
# load and tidy dataset
accel_dat <- read_csv("./original dataset/accel_data.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440, 
               names_to = "activity_minute",
               values_to = "accel_value") %>% 
  mutate(week_or_work = case_when(
    day %in% c("Saturday", "Sunday") ~ "weekend",
    day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "workday",
    TRUE ~ ""
  ),
  week = as.character(week),
  week_or_work = factor(week_or_work))
## Description on this dataset is in inline R code at the end of this chunk. week_or_work variable is binary and a factor variable with two levels


# create a table showing total activity of a day
accel_dat %>% 
  select(day_id, accel_value) %>% 
  group_by(day_id) %>% 
  summarise(activity_total_val = sum(accel_value)) %>% 
  knitr::kable()
## According to solely watching, It is hard to find a trend in it


# a graph showing 24-hour activity vs day
accel_dat %>% 
  select(day_id, accel_value) %>% 
  group_by(day_id) %>% 
  summarise(activity_total_val = sum(accel_value)) %>%
  left_join(accel_dat, by = "day_id") %>% 
  select(day_id, activity_total_val, week) %>% 
  distinct(day_id, activity_total_val, week, .keep_all = TRUE) %>% 
  ggplot(aes(x = day_id, y = activity_total_val, color = week)) +
  geom_boxplot(aes(fill = week), alpha = .5) +
  scale_x_continuous(breaks = c(0, 7, 14, 21, 28, 35)) +
  labs(
    title = "Boxplot of Acceleromator Data for Five Weeks",
    x = "Nmuber of the day",
    y = "Activity values over a day",
    caption = "Data from CUMC"
  ) 
## The trend is like a wave on a week scale and the range becomes larger, but still have a balance
  
```
The total observations in *accel_dat* is `r count(select(accel_dat, accel_value))`. The variables in this data frame is `r names(accel_dat)`. The structure of this data frame is `r dim(accel_dat)`.
